\chapter{Introduction}
Environmental monitoring is used to ensure water and air pollution levels are in
compliance with governmental regulations \parencite[i.\,e.][]{Anonymous:1996ui}, 
to monitor ozone concentrations and climate change, or for surveillance of 
industrial facilities for leakages of pollutants to just name a few 
applications.

In many of these scenarios it is feasible to have a static sensor network.  
Therefore, it is not surprising that research on optimal sensor placement at 
fixed locations exists \parencite[e.\,g.][]{Osborne:2008hi, Guestrin:2005cq, 
    Wang:kz}.  However, better results might be obtainable using mobile robots 
which can move to interesting areas and acquire more precise data there.  
Moreover, in some scenarios like disaster response, where timely information is 
needed, it might not be possible to first deploy an extensive sensor network. In 
this case mobile robots allow here to quickly identify the interesting regions.  
The problem of autonomously choosing the best locations for data acquisition is 
known as \newterm{active learning}. In the setting of environmental surveillance 
\textcite{Marchant:2012wb} also used the term \newtermAbbrev{intelligent 
    environmental monitoring}{IEM}.

In this work, I focus on a scenario proposed as part of the CompLACS project in 
\textcite{denardi2013rn}: One or more sources emit a gaseous substance or 
aerosol which is dispersed by a constant wind. The resulting plume distribution 
has to be estimated with autonomously controlled \pabbrev{unmanned aerial 
    vehicle}{UAV}. The rather steep concentration gradients and small spatial 
extent orthogonal to the main dispersion axis add to the difficulty of this 
problem.  With a few UAVs it is not possible to cover the whole volume of 
investigation using a regular pattern densely in a timely manner.  It is 
necessary to focus on measurements in specific areas.  Furthermore, measurement 
noise has to be taken into consideration.

In previous works swarms of robots have been used to localize the source of 
a plume \parencite{Jatmiko:2007df, Zarzhitsky:2005tz}. These approaches, 
however, do not allow the usage of only one robot and do not provide one with an 
estimation of the overall plume distribution. Such an estimation might be 
important for various reasons such as determining areas in which a threshold is 
exceeded or a contamination occurred. As \textcite{Reggente:2009ti} noted, 
though one could try to model the actual fluid dynamics to obtain this 
information, such computational fluid dynamics models become intractable for 
real world applications with inaccurate data.  Instead they propose to build 
a statistical model with the gas concentration measurements as random variables.

A widely used statistical model for spatial or spatio-temporal data are 
\newterm{Gaussian processes}\footnote{In geospatial statistics the modeling 
    with Gaussian processes is also known as \newterm{kriging}.}. In several 
works \parencite[e.\,g.][]{Stachniss:2008vz, Marchant:2012wb} the modeled data 
were actually gas concentrations. Also, there exists some prior work on actively 
selecting the sampling locations. \Textcite{Stranders:2008wl} do this for 
discrete locations; \textcite{Singh:2010wt} and \textcite{Marchant:2012wb} for 
continuous locations. However, none of these approaches is optimal for plume 
dispersions. This work will porpose and evaluate the improved PDUCB method for 
the given task.

The thesis is organized as follows. First, a description of the plume modeling 
scenarios will be given establishing the background of this work.  
Chapter~\ref{sec:gp} will give a general introduction into Gaussian Processes 
and discusses some topics specifically related to the modeling of plume 
distributions including online updates and active learning. Following in 
Chapter~\ref{sec:error} a number of error measures will be introduced needed to 
evaluate different approaches. Some further details on how the algorithms were 
implemented are given in Chapter~\ref{sec:tech}. The results of a number of 
simulation experiments are presented in Chapter~\ref{sec:exp}, before providing 
a short outlook on modeling time-varying plume distributions in 
Chapter~\ref{sec:timevarying}. Finally, a conclusion will be given.
